Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-08 code.tex
Updates from ShareLaTeX

\begin{figure*} 
~\hrule~
\begin{minipage}{.45\linewidth}
\small

~\\

{\bf The Line class (as described in \tion{mm}:}
Note that this is the primary data structure of HOW.
Training data is converted into a set of {\tt Line}s,
each of which is defined by a $X,Y$ pair, where $X$
has better scores than $Y$.

\begin{python}[left]
class Line:
  "Lines connect a worse cluster Y to a better one X"
  all   = []    # where to store all the lines
  using = []    # what columns to use
  F     = 0.5   # magnitude control parameter
  B     = 50    # ratio of how many columns to displace

 def __init__(i,x,y):
    if x.score < y.score:
      x,y = y,x      # x must have best score
    i.c = dist(x,y) 
    i.obj = len(x) 
    Line.all += [i]  # remember this line

  def displace(i,z):
    for j,(xj,zj) in enumerate(zip(x,z)):
      if j in Line.using and not j == i.obj:  
        z[j] = zj + Line.F*( xj - zj)
    return z

@@

{\bf Main loop, of the  HOW  instance-based planner:}
Testing data is {\em displaced}
towards   the $X$ point of the nearest {\tt Line}.
\begin{python}[left]
def HOW(training, testing):
  def train(data):
    Line.using = bPruning(data)
    clusters   = cluster(data,len(data)):
    for cluster in clusters:
      one = exemplar(cluster)
      two = exemplar(nearest(one, clusters))
      Line(one,two)

@@

   def bPruning(data,  b = Line.B):
    "Return the indexes of the 'b' best columns" 
    enough  = int(len(weights)*b/100)
    return featureWeighting(data)[:enough]

  def nearestLine(z):
    out,least=None,10000
    for line in Line.all:
      d = line.dist(z)
      if d < least: out,least = line,d
    return out

  #---------------------------------
  # begin main code for  'HOW'
  train(training)
  return [nearestLine(z).displace(z) for z in testing]
\end{python}

@@

{\bf Code for geometry (as described in \fig{where}c):}

\begin{python}[left]
def dist(x,y):
   "See Aha et al. reference [22]. XXX"

def geometry(x, y, c, z): 
  a = dist(z,x)
  b = dist(z,y) 
  x= $(a^2 + c^2 - b^2)/(2c)$ 
  y=  $\sqrt{(a^2 - max(x,0))^2}$
  return x,y

@@

def furthest(x,data, best= -1, 
                     better=lambda j,k:j > k):  
  out = None
  for one in data:
    d = dist(one,x)
    if gt(d,best): out, best = y, d
  return out
\end{python}
 \end{minipage}~~~~~~~~~\begin{minipage}{.45\linewidth} 
\small

~\\

{\bf Clustering (as described in \fig{where}a):}
\begin{python}[right]
def cluster(data, n,lvl=100):
  def splitAcross2Points(data): 
    tmp = random.choose(data)
    x = furthest(tmp, data)
    y = furthest(x, data) 
    c = dist(x,y)  
    if x.scores < y.scores:
      x,y = y,x 
    for one in data.members: 
      one.pos = geometry(x,y,c,one)
    data = sorted(data) # sorted by 'pos.x'
    return x, y, split(data)

@@

  # --------------------------------------
  # begin main code for  'cluster'
  if lvl < 1: 
     return data # stop if out of levels
  leafs = [] # Empty Set
  x,y,left,right = splitAcross2Points(data) 
  if len(left) > $\sqrt{n}$:  
     leafs += cluster(left, n, lvl - 1)  
  if len(right) > $\sqrt{n}$:  
     leafs += cluster(right,n,  lvl - 1) 
  data.has = dict(x=x,y=y,
                 l=left,r=right,
                 leafs=leafs)
  return leafs

\end{python}

@@

\begin{python}[right]
def featureWeighting(cols):
  "Returns col indexes, sorted by their weight."
  class Stats():  
    """Utility class. Handles incremental update of
       n, mean, standard dev."""
    def __init__(i,inits=[]):
      i.n = i.mu = i.m2 = 0.0
      map(i.__add__,inits)
    def sd(i) :  return (max(0.0,i.m2)/(i.n - 1))**0.5
    def __add__(i,x):
      i.n  += 1
      delta = x - i.mu
      i.mu += delta/(1.0*i.n)
      i.m2 += delta*(x - i.mu) 
    def __sub__(i,x):
      i.n  -= 1
      delta = x - i.mu
      i.mu -= delta/(1.0*i.n)
      i.m2 -= delta*(x - i.mu)

@@

  #---------------------------------
  # begin main code for  'featureWeighting'
  obj = len(cols[0]) - 1 
  return map(lambda k:k[1],
             sorted(weight1(col,i,obj) 
                    for i,col in enumerate(cols)
                    if not i == obj))
\end{python}
\end{minipage}
~\hrule~
\caption{HOW (Python-style psuedo-code).
For brevity's sake, this code skips certain low-level details.
For a full working implementation, see https://github.com/ai-se/HOW1/src.
Functions shown in \textcolor{blue}{{\bf blue}} are defined somewhere in this figure.}\label{fig:howcode}   
\end{figure*}

>>>>

\begin{figure*} 
~\hrule~
\begin{minipage}[t]{.45\linewidth}
\scriptsize\vspace{1mm}
{\bf The Line class (as described in \tion{mm}):}

\begin{python}[left]
class Line:
  """The central data structure of HOW. Training data 
  is converted into a set of Lines, each of which is 
  defined by a 'X,Y' pair, where 'X'
  has better scores than 'Y'."""
  all   = []    # where to store all the lines
  weights=[]    # (weight,feature) pairs, sorted
  using = []    # what columns to use
  F     = 0.5   # magnitude control parameter
  B     = 50    # ratio of how many columns to displace

 def __init__(i,x,y):
    if x.score < y.score:
      x,y = y,x      # x must have best score
    i.x, i.y, i.c = x, y, dist(x,y) 
    Line.all += [i]  # remember this line

  def displace(i,z):
    for j in Line.using:
      z[j] = z[j] + Line.F*(x[j] - z[j])
    return z

@@

{\bf Main loop, of the  HOW  instance-based planner:}

\begin{python}[left]
def HOW(training, testing, b = Line.b):
  """ Testing data is displaced towards
  the X point of the nearest Line."""
  def train(data):
    Line.weights= weightedFeatures(data)
    b           = int(len(Line.weights)*B/100)
    Lines.using = Line.weights[:b])
    clusters    = [k for k in 
                   leafs(cluster(data,len(data)))]
    for cluster in clusters:
      one = exemplar(cluster)
      two = exemplar(nearest(one, clusters))
      Line(one,two)

@@

  def nearestLine(z):
    out,least=None,10000
    for line in Line.all:
      d = line.dist(z)
      if d < least: out,least = line,d
    return out
  # ---- begin main code for  'HOW' ---------------
  train(training)
  return [nearestLine(z).displace(z) for z in testing]
\end{python}

@@

{\bf Misc utilities; e.g. code for geometry (as described in \fig{where}c):}

\begin{python}[left]
def geometry(x, y, c, z): 
  a = dist(z,x)
  b = dist(z,y) 
  x= $(a^2 + c^2 - b^2)/(2c)$ 
  y=  $\sqrt{(a^2 - max(x,0))^2}$
  return x,y

def dist(x,y):
   tmp = sum(w*(x[j] - y[j])**2 for w,j in Line.using) 
   return tmp**0.5 / len(Line.using)**2

@@

def furthest(x,data, best= -1, 
                     better=lambda j,k:j > k):  
  out = None
  for one in data:
    d = dist(one,x)
    if gt(d,best): out, best = y, d
  return out

def leafs(data):
  leaf = True
  for more in data.leafs:
    if more:
       leaf = False
       for one in leafs(more):
         yield one
  if leaf: 
    yield data
\end{python}
 \end{minipage}~~~~~~~~~\begin{minipage}[t]{.45\linewidth} 
\scriptsize\vspace{1mm}
{\bf Clustering (as described in \fig{where}a):}
\begin{python}[right]
def cluster(data, n,lvl=100):

  def splitAcross2Points(data): 
    tmp = random.choose(data)
    x = furthest(tmp, data)
    y = furthest(x, data) 
    c = dist(x,y)  
    if x.scores < y.scores:
      x,y = y,x 
    for one in data.members: 
      one.pos = geometry(x,y,c,one)
    data = sorted(data) # sorted by 'pos.x'
    return x, y, split(data)

@@

  # --- begin main code for  'cluster' --------
  if lvl < 1: 
     return data # stop if out of levels
  leafs = [] # Empty Set
  x,y,left,right = splitAcross2Points(data) 
  if len(left) > $\sqrt{n}$:  
     leafs += cluster(left, n, lvl - 1)  
  if len(right) > $\sqrt{n}$:  
     leafs += cluster(right,n,  lvl - 1) 
  date.leafs = leafs
  return leafs
\end{python}

@@

\begin{python}[right]
def weightedFeatures(cols):
  "Returns col indexes, sorted by their weight."

  class Stats():  
    """Utility class. Handles incremental update of
       n, mean, standard dev."""
    def __init__(i,inits=[]):
      i.n = i.mu = i.m2 = 0.0
      map(i.__add__,inits)

    def sd(i) :  
      return (max(0.0,i.m2)/(i.n - 1))**0.5

    def __add__(i,x):
      i.n  += 1
      delta = x - i.mu
      i.mu += delta/(1.0*i.n)
      i.m2 += delta*(x - i.mu)

    def __sub__(i,x):
      i.n  -= 1
      delta = x - i.mu
      i.mu -= delta/(1.0*i.n)
      i.m2 -= delta*(x - i.mu)

@@

  # ---- begin main code for 'featureWeighting' -----
  obj = len(cols[0]) - 1 
  return sorted(weight1(col,i,obj) 
                    for i,col in enumerate(cols)
                    if not i == obj))
\end{python}
\end{minipage}
~\hrule~
\caption{HOW (Python-style psuedo-code).
For brevity's sake, this code skips certain low-level details.
For a full working implementation, see https://github.com/ai-se/HOW1/src.
Functions shown in \textcolor{blue}{{\bf blue}} are defined somewhere in this figure.}\label{fig:howcode}   
\end{figure*}

### 2015-05-08 how.tex
Updates from ShareLaTeX

\lstnewenvironment{python}[1][]{
\lstset{
mathescape,
numbers=right,
numberstyle=\scriptsize,
stepnumber=1,
numbersep=0.5em,
xleftmargin=0em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\scriptsize\setstretch{0.8},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap, Line,bPruning,clister,train, exemplar,nearestLine,dist,displace,geometry,splitAcross2Points,How,nearest,bPruning,Stats,divide,recurse,weight1,project, furthest, split, WHERE,clusterer, getContours, envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
firstnumber=last
%
}}{}

@@

HERE's feature weighting algorithm
comes from the CART regression tree learner~\cite{Breiman1984}.
It sorts independent variables
 according to how well they can reduce the variance
of a  numeric objective.
If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
then the expected
value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
where $v$ is the standard deviation of the objective.
A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
then split at $\argmin_i w_i$, then recursing into each half.  For more details, see  \fig{howcode}, line~109.

@@

 \begin{figure*}[htbp!]
\centering
\begin{minipage}{0.30\linewidth}
\includegraphics[width=\linewidth]{_figs/Apache.pdf}
\end{minipage}
\begin{minipage}{0.30\linewidth}
\includegraphics[width=\linewidth]{_figs/BDBJ.pdf}
\end{minipage}
\begin{minipage}{0.30\linewidth}
\includegraphics[width=\linewidth]{_figs/BDBC.pdf}
\end{minipage}

@@

{\em 3. Evaluation bias:} The issue was discussed above in \tion{assess} where
we asserted that plans can be assessed via predictors-- with the all important
priviso that the predictors are not performing poorly.

>>>>

\lstnewenvironment{python}[1][]{
\lstset{
mathescape,
numbers=right,
numberstyle=\scriptsize,
stepnumber=1,
numbersep=0.5em,
xleftmargin=0em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\scriptsize\setstretch{0.8},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap, Line,bPruning,clister,train,leafs,weightedFeatures, exemplar,nearestLine,dist,displace,geometry,splitAcross2Points,leaves,How,nearest,bPruning,Stats,divide,recurse,weight1,project, furthest, split, WHERE,clusterer, getContours, envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
firstnumber=last
%
}}{}

@@

HERE's feature weighting algorithm
comes from the CART regression tree learner~\cite{Breiman1984}.
It sorts independent variables
 according to how well they can reduce the variability
of a  numeric objective.
If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
then the expected
value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
where $v$ is the standard deviation of the objective.
A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
then split at $\argmin_i w_i$, then recursing into each half.  For more details, see  \fig{howcode}, line~109.

@@

 \begin{figure*}[htbp!]
\centering
\begin{minipage}{0.30\linewidth}
\includegraphics[width=\linewidth]{_figs/Apache.pdf}
\end{minipage}
\begin{minipage}{0.30\linewidth}
\includegraphics[width=\linewidth]{_figs/BDBJ.pdf}
\end{minipage}
\begin{minipage}{0.30\linewidth}
\includegraphics[width=\linewidth]{_figs/BDBC.pdf}
\end{minipage}\\[0.25cm]

@@

{\em 3. Evaluation bias:} The issue was discussed above in \tion{assess} where
we asserted that plans can be assessed via predictors-- with the all important
proviso that the planning should {\em not} be attempted if the predictors
are performing poorly.
